---
title: "Diamonda dataset analysis"
output:
  html_document:
    df_print: paged
---

 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
library(tidyverse)
library(DT)
library(ggplot2)
options(knitr.table.format = "html")

#diamonds data reading 
jems <- read.csv("C:/Users/USER/Desktop/DS_EDA/DS_EDA_Daimonds/diamonds.csv")
```
\# data summary 
```{r}
jems %>% 
  summary()%>% 
  datatable()
```
\# price average - median
```{r}
jems %>% 
  summarise(avg= mean(price), median=median(price))
```
\# price avg & median deepening on the clarity type 

```{r}
jems %>% 
  group_by(clarity) %>% 
  summarise(avg = mean(price), median=median(price))
```
\# price avg & median deepening on the cut type 
```{r}
jems %>% 
  group_by(cut) %>% 
  summarise(avg = mean(price), median=median(price))
```
\# price avg & median deepening on the color type 
```{r}

jems %>% 
  group_by(color) %>% 
  summarise(avg = mean(price), median=median(price))
```
I wanted to play around and see if i can predict the carat from the size using very basic formula so i did the following:

diamonds Weight in carats = Length x Width x Depth x Coefficient (estimated because it depends about the cut shape)

L/W Ratio 1.00: Coefficient = 0.0060

L/W Ratio 1.25: Coefficient = 0.0080

L/W Ratio 1.50: Coefficient = 0.0090

L/W Ratio 2.00: Coefficient = 0.0100

L/W Ratio 2.50: Coefficient = 0.0105

since there is only about 10 diamonds that they are far from Coefficient 1 we will use the 0.006

```{r}
jems <- jems %>% 
  mutate(ratio = x / y)
         
jems <- jems %>% 
  mutate(carats = x * y * z  * 0.006)

# comparing my results 
jems <- jems %>% 
  mutate( diff_percenatge= abs(((carat-carats)/carat)*100))

# avg difference 
jems %>% 
  filter(diff_percenatge > 0 & diff_percenatge < 100 ) %>% 
  summarise(avg = mean(diff_percenatge), median= median(diff_percenatge)) %>% 
  datatable()
```
Plotting the results 
```{r}
jems %>% 
  filter(diff_percenatge > 0 & diff_percenatge < 100 ) %>% 
  ggplot(aes( x= carat,y=carats))+ geom_jitter()+
  geom_smooth(method = "lm", se = FALSE)+
  xlab("Actual")+
  ylab("Predectied")
```
pretty good results :)  

price error from the prediction based on the original price of the diamonds

```{r}
jems <- jems %>% 
  mutate(price_unit= price/carat)
         
jems <- jems %>% 
  mutate(new_price= price_unit*carats)

jems <- jems %>% 
  mutate(price_diff= abs(price - new_price))
  
jems %>% 
  filter(diff_percenatge > 0 & diff_percenatge < 100 ) %>% 
  summarise(avg = mean(price_diff), median= median(price_diff)) %>% 
              datatable()
```

